Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by \textsc{DigitalSnow} ANR-11-BS02-009, \textsc{KIDICO} ANR-2010-BLAN-0205 and \textsc{PRIMES Labex} ANR-11-LABX-0063/ ANR-11-IDEX-0007 research grants.}

### 2015-10-23 latex/article.tex
minor modifications

Volumetric objects are being more and more popular in many
applications ranging from object modeling and rendering in Computer
Graphics to geometry processing in Medical Imaging or Material
Sciences. When considering large volumetric data, interactive
visualization of such images (or isovalue surfaces) is a complex
problem. Such issues become even more difficult when dynamic
volumetric datasets are considered. Beside such isosurface
visualization, we are also interested in performing geometry
processing on the digital object and to explore different parameter
settings of the geometry processing tool.  In this work, we focus on
interactive visualization of volumetric data isosurfaces combined with
a curvature tensor (mean/Gaussian curvature, principal curvatures
directions, \ldots) estimation at each point of the digital
surface. Curvature maps on a given isosurface are key information when
exploring 3D structures in Material Sciences \todo{REF}. Furthermore,
most curvature estimators require a parameter fixing the scale
to which the computation is performed. For short, such parameter
(integration radius, convolution kernel size, \ldots) specifies noise
level on the object surface. As a consequence, exploring interactively
different values of such parameter is mandatory.
\vspace{0.2cm}

@@

\sloppy\noindent\textbf{Related works}\quad Extracting and Visualizing
isosurface on volumetric data has been widely investigated since the
seminal Marching Cubes approach by \textsc{Lorensen} and
\textsc{Cline}~\cite{lorensen1987marching}. This approach being data parallel,
GPU implementation of this method is very efficient
\cite{tatarchuk2007real}. However, such technique generates a lot of
triangles which is not well suited to large digital data. Hence,
adaptive approaches have been proposed in order to optimize the
triangular mesh resolution to the object geometry or depending on the
viewpoint. In this topic, many solutions have been developed in
Computer
Graphics~\cite{shu1995adaptive,schaefer2004dual,lengyel2010voxel,DBLP:journals/cgf/LewinerMPPL10,DBLP:journals/cvgip/LobelloDD14}.
The method developed by \textsc{Lengyel} \textit{et al.}
\cite{lengyel2010voxel} is the best suited to our needs.  However, it
was only developed on CPU to visualize in real-time static or near
static data, combining octree with new Marching Cubes configurations
to generate adaptive meshes. In this paper we propose a GPU pipeline
inspired from this method.  It leaves the CPU free for other tasks and
avoids synchronization with the GPU.  Our high framerate also allows
this method to run in real-time on fully dynamic data.  The
triangulation is controlled by a GPU LoD criterion that ensures that
triangles project onto more than one pixel, thus removing geometrical
aliasing.

@@

\sloppy In our context, we consider digital shapes (any subset $Z$ of $\Z^d$)
and boundaries of digital shapes $\DigBoundary{Z}$. We denote by $\DSh$ the
Gauss digitization of $\Shape$ in a $d-$dimensional grid with grid step $h$,
\emph{i.e.}  $\DSh \EqDef \{ \vz \in \Z^d,\, h\cdot\vz\in\Shape\}$. For such
digitized set $\DigShape\EqDef\DSh$, $\Body{\DigShape}{h}$ is a subset of $\R^d$
corresponding to the union of hypercubes centered at $h\cdot \vz$ for $\vz\in Z$
with edge length $h$. By doing so, both $\partial\Shape$ and
$\partial\Body{Z}{h}$ are topological boundary of objects lying in the same
space (See Fig. \ref{fig:notations}-b). Note that the combinatorial digital
boundary $\DigBoundary{Z}$ of $Z$ made of cells in a cellular Cartesian complex
(\emph{pointels}, \emph{linels}, \emph{surfels}, \ldots), can be trivally
embedded into $\R^d$ such that it coincides with $\partial\Body{Z}{h}$.

@@

In this section, we detail the adaptive isosurface extraction
algorithm. The proposed approach is based on an octree representation
of the input object on which an efficient adaptive triangulation
of the isosurface is constructed on GPU. Such approach is motivated by
the fact that the hierarchical representation of the object allows us
to handle large datasets and to locally adapt the level of details
\wrt the geometry or camera position. We first present the
octree representation and then the isosurface extraction from
hierarchical octree cells.

@@

Representing a hierachical structure on GPU is usually challenging since the GPU
memory is limited and non consistent %\todo{fix coherent/aligned/coalesced?}
memory access (\emph{e.g.} tree traversal using pointer dereferencing) may lead
to huge payload from cache misses.  Efficient spatial tree encoding can be
achieved using pointerless structures such as linear quadtrees or octrees, see
for instance \textsc{Gargantini}~\cite{gargantini1982effective}. This structure
consists in indexing each cell by a \emph{Morton code}: the code of children
cells are defined by the code of the parent suffixed by two bits (in dimension
2, three bits in dimension 3) (see Figure
\ref{fig_quadtree_partitionning}-\emph{left}). A cell's code encodes its
position \wrt its parent cell and its complete path to the tree root. Hence,
whatever their depth, all cell codes are simply stored in a linear vector on
memory. Furthermore, cell operations such as subdivision, merging, neighbors
access can be efficiently implemented using bitwise operations on the Morton
code. In the following, we use the GPU implementation proposed by \textsc{Dupuy}
\textit{et al.}~\cite{dupuy2014quadtrees}. By using their approach, Morton codes
and bitwise operations are optimized to match with GPU hardware requirements.

@@

Using this spatial data structure, triangulated mesh can be constructed
from a Marching Cubes approach \cite{lorensen1987marching} (MC for short): the
triangulation is generated from local triangle patches computed on
local cell configurations. Since triangle patches are locally
computed, such approach is fully parallel and easy to implement on GPU
hardwares. However, since adjacent cells may not have the same depth
on the octree, original \textsc{Lorensen} and \textsc{Cline}'s rules need to be
updated (see Figure~\ref{fig_quadtree_partitionning}-\emph{right}). Many authors have addressed this problem both for primal and
dual meshes
\cite{shu1995adaptive,schaefer2004dual,lengyel2010voxel,DBLP:journals/cgf/LewinerMPPL10,DBLP:journals/cvgip/LobelloDD14}.

In the following, we use the extension of MC configurations to handle
adaptive structures proposed by \textsc{Lengyel} \emph{et al.}~\cite{lengyel2010voxel}. First, this approach constrains the octree
structure in order to make sure that the depth difference between any
two adjacent cells is at most one. Then, \textsc{Lengyel} \emph{et al.} propose
specific MC configurations for such transition cells,  leading to a
crack-free triangulation.

@@

As illustrated in Figure~\ref{fig_quadtree_partitionning}-\emph{middle}, we
propose a viewpoint dependent criterion to decide if a cell needs to be refined:
the closer we are to the camera, the finer the cells are. In addition to the
distance criterion, we define an angular criterion to prevent refining cells
outside the camera frustrum. Using a criterion based on the distance between a
cell and the camera is well suited to GPU since it can be evaluated
independently on all cells. Hence, we can efficiently determine if a cell must
be subdivided or not. Figure~\ref{fig_lod_octree} illustrates the level of
details (LoD for short) criterion we use. More precisely, in dimension 2, if
$\alpha$ denotes the viewing angle, an object at distance $d$ from the camera
with diameter greater than $2\cdot d\cdot\tan(\alpha)$ is projected into more
than one pixel (see Figure~\ref{fig_lod_octree}-\emph{left}). Hence, the
distance criterion is based on the ratio (\emph{visibility ratio} in the
following) between the cell diameter $l(c)$ (power of 2 depending on the depth),
and $2\cdot d_c\cdot\tan(\alpha)$, $d_c$ being the distance between the cell
center and the camera. For a given cell $c$, split and merge decision are based
on this visibility ratio:
%
\begin{itemize}
  \item $c$ is split if its children
cells have a visibility ratio greater than  constant $k$\,;
  \item $c$ and its sibling cells are merged if their parent cell $c'$ has a visibility ratio lower than $k$\;
  \item otherwise, the cell $c$ stays for the next frame.
\end{itemize}
%
\begin{figure}[!h]
  \begin{center}
    \begin{overpic}[width=3.5cm]{figs/criterion}
      \put(33,40){$\alpha$}
      \put(50,45){$d_c$}
      \put(95,65){$l(c)$}
    \end{overpic}~~~~~
    \includegraphics[width=3.5cm]{viewlod2_small}~~~~~
    \includegraphics[width=3cm]{subdivision}
  \end{center}
  \caption{Notations \emph{(left)} and adaptive meshing in dimension 2 using the LoD distance and angular criterion \emph{(middle)}. Based on LoD criterion
    evaluation, the cell buffer is updated between each frame \emph{(right)}.}
  \label{fig_lod_octree}
\end{figure}
%
In addition to the distance criterion, we add an angular term in order
to prevent subdivision of cells far from the camera frustrum (see
Figure~\ref{fig_lod_octree}-\emph{middle}). We do not go into the
details but please note that stay, split and merge decisions of all cells are computed in
parallel from all cell morton codes.

@@

\begin{figure}
  \begin{center}
    {\includegraphics[width=2.7cm]{figs/Octa_mean}}
    {\includegraphics[width=2.7cm]{figs/Octa_gaussian}}
    {\includegraphics[width=2.7cm]{figs/Octa_k1}}
    {\includegraphics[width=2.7cm]{figs/Octa_k2}}\\
    {\includegraphics[width=2.7cm]{figs/Octa_dir_min}}
    {\includegraphics[width=2.7cm]{figs/Octa_dir_max}}
    {\includegraphics[height=2.0cm]{figs/Octa_dir_min_zoom}}
    {\includegraphics[height=2.0cm]{figs/Octa_dir_max_zoom}}\\
    % {\includegraphics[height=4.0cm]{figs/EG_dragon_normal_zoom}}
    {\includegraphics[width=2.2cm]{figs/function_mean_0}}
    {\includegraphics[width=2.2cm]{figs/function_mean_1}}
    {\includegraphics[width=2.2cm]{figs/function_mean_2}}
    {\includegraphics[width=2.2cm]{figs/function_mean_3}}
    {\includegraphics[width=2.2cm]{figs/function_mean_4}}
  \end{center}
  \caption{\emph{First row:} Mean, Gaussian, first and second principal curvature estimation on an OctaFlower with a digital domain of $130^3$.
  \emph{Second row:} First and second principal directions estimation (and zoom) on an OctaFlower.
  \emph{Third row:} Mean curvature computed in real-time on a dynamic object.}
  \label{fig:full}
\end{figure}

We first evaluate the curvature estimation on a full-resolution geometry
obtained by disabling the LoD criterion at the mesh generation step.
Figure~\ref{fig:full}-top and middle shows results of curvature tensor
estimation (mean, Gaussian, first and second principal curvatures, first and
second principal directions) on an OctaFlower at level $l_0$ (ground truth).
Figure~\ref{fig:full}-bottom shows mean curvature estimation in real-time on a
dynamic object. For this case, we simply evaluate the implicit expression at
each vertex of a cell to decide if such cell is included or not into
$\Ball{R}{\vx}\cap \Shape$ instead of updating the \emph{mipmap texture} of
densities at each time step. For all illustrations, we use directly normal
vectors computed by algorithm discussed in Section~\ref{sec:curv-tens-estim}.

@@

  \end{center}
  \caption{Comparison of number of texture fetches, $L_2$ and $L_\infty$ error obtained on OctaFlower with a digital domain of $130^3$. We measured it using radii of 8 and 16.
  The object was triangulated at full resolution with a regular grid and contains 282,396 vertices. The timings
    only consider the curvature estimation step.\label{tab:full-res-stat}}
\end{table}

@@

\begin{figure}[!htbp]
  \begin{center}
    \begin{tikzpicture}[scale=0.7]
      \begin{axis}[
          ybar stacked,
          ymode = log,
          ymax = 6000,
	  	  bar width=20pt,
	  	  nodes near coords,
	  	  every node near coord/.style={ font=\tiny, inner sep=0pt},
          enlargelimits=0.15,
          ylabel={\#Time (in ms)},
          symbolic x coords={Octa-R8, Octa-R16, Snow-R10, Snow-R20,
	      Dragon-R8, Dragon-R20},
          xtick=data,
          point meta=rawy,
          x tick label style={rotate=45,anchor=east},
        ]
        %Geometry Generation
        \addplot+[ybar, color=black, fill=cgeometry] plot coordinates {(Octa-R8,17.283000) (Octa-R16,18.458000)
          (Snow-R10,9.831000) (Snow-R20,9.653000)  (Dragon-R8,136.300000) (Dragon-R20,145.450000)};
        %l4
        \addplot+[ybar, color=black, fill=clquatre] plot coordinates {(Octa-R8,1) (Octa-R16,2.509000)
          (Snow-R10,6.085000) (Snow-R20,9.445000)  (Dragon-R8,3.389000) (Dragon-R20,20.228000)};
        %l3
        \addplot+[ybar, color=black, fill=cltrois] plot coordinates {(Octa-R8,2.527000) (Octa-R16,4.159000)
          (Snow-R10,10.629000) (Snow-R20,11.542000) (Dragon-R8,20.397000) (Dragon-R20,26.710000)};
        %l2
        \addplot+[ybar, color=black, fill=cldeux] plot coordinates {(Octa-R8,3.783000) (Octa-R16,9.006000)
          (Snow-R10,13.777000) (Snow-R20,37.700000) (Dragon-R8,29.967000) (Dragon-R20,92.435000)};
        %l1
        \addplot+[ybar, color=black, fill=clun] plot coordinates {(Octa-R8, 11.004000) (Octa-R16,38.687000)
          (Snow-R10,39.627000) (Snow-R20,217.954000) (Dragon-R8,59.840000) (Dragon-R20,506.549000)};
        %l0
        \addplot+[ybar, color=black, fill=clzero] plot coordinates {(Octa-R8, 37.273000) (Octa-R16,244.353000)
          (Snow-R10,236.194000) (Snow-R20,1655.680000) (Dragon-R8,322.674000) (Dragon-R20,3763.422000)};
        %\legend{\strut Geometry, \strut Curvature, \strut xx, \strut xx}
      \end{axis}
	  \end{tikzpicture}
%
    \begin{tikzpicture}[scale=0.7]
      \begin{axis}[
          ybar stacked,
          ymax = 6000,
          ymode = log,
	  	  bar width=20pt,
	  	  nodes near coords,
	  	  every node near coord/.style={ at={(0,0)}, font=\tiny, inner sep=0pt},
          enlargelimits=0.15,
          %ylabel={\#Time (in ms)},
          symbolic x coords={Octa-R8, Octa-R16, Snow-R10, Snow-R20,
	      Dragon-R8, Dragon-R20},
          xtick=data,
          point meta=rawy,
          x tick label style={rotate=45,anchor=east},
        ]
        %Geometry Generation
        \addplot+[ybar,color=black, fill=cgeometry] plot coordinates {(Octa-R8,17.283000) (Octa-R16,17.183000)
          (Snow-R10,9.383000) (Snow-R20,9.767000)  (Dragon-R8,127.347000) (Dragon-R20,127.166000)};
        %Hierarchical measurement
        \addplot+[ybar, color=black, fill=clzero] plot coordinates {(Octa-R8,18.095000) (Octa-R16,64.186000)
          (Snow-R10,96.747000) (Snow-R20,364.547000)  (Dragon-R8,147.662000) (Dragon-R20,787.459000)};
        %\legend{\strut Geometry, \strut Curvature, \strut xx, \strut xx}
      \end{axis}
     \end{tikzpicture}
     \end{center}
  %
  \caption{Timings in milliseconds (in logscale) obtained while visualizing an adaptive
  triangulation on three objects -- OctaFlower (digital domain of $130^3$ ), Snow
  Microstructures ($233^3$) and XYZ-Dragon ($510^3$) --  by computing the
  curvature with a regular grid (\emph{left}) and with hierarchical
  algorithm (\emph{right}), with two different radii for each object.
  %
  In \emph{orange color}: time required to extract the triangulation.
  %
  In \emph{blue color}: time required to compute the curvature tensor at
  different levels: from $l_4$ (light blue) to $l_0$ (dark blue).
  %
  Timings are given using a NVIDIA GeForce GTX 850M GPU.
  }
  %
  \label{fig:timings}
\end{figure}

When dealing with large datasets, we can not expect real-time curvature tensor
estimation due to the huge amount of data to process. By computing an
dynamically refined approximate curvature tensor estimation joined with an
adaptive triangulation (as discussed in Section~\ref{sec:hierarchical}), we
manage to maintain a real time framerate by giving control over the amount of
data to process at each frame. In Figure~\ref{fig:timings}, we compare timings
(in logarithmic scale) when the triangulation is dynamically refining according
to its distance to the camera. We measured those timings using the regular and
the hierarchical algorithm.

We can first note that the required time for an object to compute ground truth
curvature is usually as high as the time required to extract the geometry and
compute all the above levels. Using approximations is thus mandatory, since for
big objects such as the XYZ-Dragon ($510^3$) with big radii the approximate
computation is faster than the $l_0$ while results are enough close to the
$l_0$. For example, with $R=20$, the approximated visualization at $l_2$ takes
only 300ms while the $l_0$ visualization takes more than 4 seconds.

We can thus get an approximated visualisation of our object as soon as we run
the application, allowing for real time interactions with the object (changing
the visualized value, curvature radius, \ldots). It is also clear on those
graphs that using a hierarchical decomposition reduces the curvature computation
time. For the XYZ-Dragon with $R=20$, it needs almost 4 seconds to compute the
$l_0$ level, whereas using the hierarchical algorithm, it only needs 700ms.
However, due to the precomputation, this algorithm is biased and creates an
error (presented in Table~\ref{tab:full-res-stat}) that needs to be considered.

@@

Figure~\ref{fig:adaptive} shows curvature computation and exploration in
real-time on large datasets: XYZ-Dragon (with digital domain of $512^3$) and
Snow microstructures ($233^3$).

>>>>

Volumetric objects are being more and more popular in many
applications ranging from object modeling and rendering in Computer
Graphics to geometry processing in Medical Imaging or Material
Sciences. When considering large volumetric data, interactive
%visualization of such images (or isovalue surfaces) is a complex
visualization of those objects (or isosurfaces) is a complex
problem. Such issues become even more difficult when dynamic
%volumetric datasets are considered. Beside such isosurface
volumetric datasets are considered. Beside
visualization, we are also interested in performing geometry
processing on the digital object and to explore different parameter
settings of the geometry processing tool.  In this work, we focus on
interactive visualization of volumetric data isosurfaces combined with
a curvature tensor (mean/Gaussian curvature, principal curvatures
directions, \ldots) estimation at each point of the digital
surface. 
%Curvature maps on a given isosurface are key information when
%exploring 3D structures in Material Sciences \todo{REF}. 
Furthermore,
most curvature estimators require a parameter fixing the scale
to which the computation is performed. For short, such parameter
(integration radius, convolution kernel size, \ldots) specifies noise
level on the object surface. As a consequence, exploring interactively
different values of such parameter is mandatory.
\vspace{0.2cm}

@@

\sloppy\noindent\textbf{Related works}\quad Extracting and Visualizing
isosurface on volumetric data has been widely investigated since the
seminal Marching Cubes approach by \textsc{Lorensen} and
\textsc{Cline}~\cite{lorensen1987marching}. This approach being data parallel,
GPU implementation of this method is very efficient
\cite{tatarchuk2007real}. However, such technique generates a lot of
triangles which is not well suited to large digital data. Hence,
adaptive approaches have been proposed in order to optimize the
triangular mesh resolution to the object geometry or depending on the
viewpoint. In this topic, many solutions have been developed in
Computer
Graphics~\cite{shu1995adaptive,schaefer2004dual,lengyel2010voxel,DBLP:journals/cgf/LewinerMPPL10,DBLP:journals/cvgip/LobelloDD14}.
The method developed by \textsc{Lengyel} \textit{et al.}
\cite{lengyel2010voxel} is the best suited to our needs. 
%However, it
%was only developed on CPU to visualize in real-time static or near
%static data, combining octree with new Marching Cubes configurations
%to generate adaptive meshes. 
It combines an octree space partitioning with new Marching Cubes configuration
to generate adaptive meshes, but was only developed on the CPU to visualize
static or near static data.
In this paper we propose a full GPU pipeline
%inspired from this method.  
inspired from this method, using a GPU-friendly Lod criterion to maintain a view dependent triangulation.
%It leaves the CPU free for other tasks and avoids synchronization with the GPU.  
%Our high framerate also allows
Our high framerate allows
this method to run in real-time on fully dynamic data.  
%The triangulation is controlled by a GPU LoD criterion 
%that ensures that
%triangles project onto more than one pixel, thus removing geometrical
%aliasing.

@@

\sloppy In our context, we consider digital shapes (any subset $Z$ of $\Z^d$)
and boundaries of digital shapes $\DigBoundary{Z}$. We denote by $\DSh$ the
Gauss digitization of $\Shape$ in a $d-$dimensional grid with grid step $h$,
\emph{i.e.}  $\DSh \EqDef \{ \vz \in \Z^d,\, h\cdot\vz\in\Shape\}$. For such
digitized set $\DigShape\EqDef\DSh$, $\Body{\DigShape}{h}$ is a subset of $\R^d$
corresponding to the union of hypercubes centered at $h\cdot \vz$ for $\vz\in Z$
with edge length $h$. By doing so, both $\partial\Shape$ and
$\partial\Body{Z}{h}$ are topological boundary of objects lying in the same
space (See Fig. \ref{fig:notations}-b). Note that the combinatorial digital
boundary $\DigBoundary{Z}$ of $Z$ made of cells in a cellular Cartesian complex
(\emph{pointels}, \emph{linels}, \emph{surfels}, \ldots), can be trivially
embedded into $\R^d$ such that it coincides with $\partial\Body{Z}{h}$.

@@

In this section, we detail the adaptive isosurface extraction
algorithm. 
%The proposed approach is based on an octree representation
%of the input object on which an efficient adaptive triangulation
%of the isosurface is constructed on GPU. 
The proposed approch use an octree representation of the input object
on which an adaptive Marching Cube is used to build the isosurface efficiently.
Such approach is motivated by
the fact that the hierarchical representation of the object allows us
to handle large datasets and to locally adapt the level of details
\wrt the geometry or camera position. We first present the
octree representation and then the isosurface extraction from
hierarchical octree cells.

@@

Representing a hierachical structure on GPU is usually challenging since
the GPU is a data parallel component and is unable to handle recursivity.

%since the GPU
%memory is limited and non consistent %\todo{fix coherent/aligned/coalesced?}
%memory access (\emph{e.g.} tree traversal using pointer dereferencing) may lead
%to huge payload from cache misses.

Efficient spatial tree encoding can be
achieved using pointerless structures such as linear quadtrees or octrees ( see
for instance \textsc{Gargantini}~\cite{gargantini1982effective} ). This structure
%consists in indexing each cell by a \emph{Morton code}: the code of children
indexes each cell by a \emph{Morton code}: the code of children
cells are defined by the code of the parent suffixed by two bits (in dimension
2, three bits in dimension 3) (see Figure
\ref{fig_quadtree_partitionning}-\emph{left}). A cell's code encodes its
position \wrt its parent cell and its complete path to the tree root. 
%Hence, whatever their depth, all cell codes are simply stored in a linear vector on memory. 
Hence, the tree is fully represented as a linear vector of its leaves, whatever their depth.
Furthermore, cell operations such as subdivision, merging, neighbors
access can be efficiently implemented using bitwise operations on the Morton
code. In the following, we use the GPU implementation proposed by \textsc{Dupuy}
\textit{et al.}~\cite{dupuy2014quadtrees} that we extended to the third dimension. 
By using their approach, Morton codes
and bitwise operations are optimized to match with GPU hardware requirements.

@@

Using this spatial data structure, triangulated mesh can be constructed
from a Marching Cubes approach \cite{lorensen1987marching} (MC for short): the
triangulation is generated from local triangle patches computed on
local cell configurations. Since triangle patches are locally
computed, such approach is fully parallel and easy to implement on GPU
hardwares. However, since adjacent cells may not have the same depth
in the octree, original \textsc{Lorensen} and \textsc{Cline}'s rules need to be
updated (see Figure~\ref{fig_quadtree_partitionning}-\emph{right}). Many authors have addressed this problem both for primal and
dual meshes
\cite{shu1995adaptive,schaefer2004dual,lengyel2010voxel,DBLP:journals/cgf/LewinerMPPL10,DBLP:journals/cvgip/LobelloDD14}.

In the following, we use the extension of MC configurations to handle
adaptive structures proposed by \textsc{Lengyel} \emph{et al.}~\cite{lengyel2010voxel}. 
First, this approach constrains the octree
structure in order to make sure that the depth difference between any
two adjacent cells is at most one. 
Then, \textsc{Lengyel} \emph{et al.} introduce the concept of transition cells.
Those cells are defined to be inserted between two neighbouring octree cells of different depth.
With specific MC configurations for such transition cells, a crack free triangulation can
be extracted.
%Then, \textsc{Lengyel} \emph{et al.} propose
%specific MC configurations for such transition cells,  leading to a
%crack-free triangulation.

@@

As illustrated in Figure~\ref{fig_quadtree_partitionning}-\emph{middle}, we
propose a viewpoint dependent criterion to decide if a cell needs to be refined:
the closer we are to the camera, the finer the cells are. 
%In addition to the
%distance criterion, we define an angular criterion to prevent refining cells
%outside the camera frustrum. 
Using a criterion based on the distance between a
cell and the camera is well suited to GPU since it can be evaluated
independently on all cells. Hence, we can efficiently determine if a cell must
be subdivided or not. Figure~\ref{fig_lod_octree} illustrates the level of
details (LoD for short) criterion we use. More precisely, in dimension 2, if
$\alpha$ denotes the viewing angle, an object at distance $d$ from the camera
with diameter greater than $2\cdot d\cdot\tan(\alpha)$ is projected into more
than one pixel (see Figure~\ref{fig_lod_octree}-\emph{left}). Hence, the
distance criterion is based on the ratio (\emph{visibility ratio} in the
following) between the cell diameter $l(c)$ (power of 2 depending on the depth),
and $2\cdot d_c\cdot\tan(\alpha)$, $d_c$ being the distance between the cell
center and the camera. For a given cell $c$, split and merge decision are based
on this visibility ratio:
%
\begin{itemize}
  \item $c$ is split if its children
cells have a visibility ratio greater than  constant $k$\,;
  \item $c$ and its sibling cells are merged if their parent cell $c'$ has a visibility ratio lower than $k$\;
  \item otherwise, the cell $c$ stays for the next frame.
\end{itemize}
%
\begin{figure}[!h]
  \begin{center}
    \begin{overpic}[width=3.5cm]{figs/criterion}
      \put(33,40){$\alpha$}
      \put(50,45){$d_c$}
      \put(95,65){$l(c)$}
    \end{overpic}~~~~~
    \includegraphics[width=3.5cm]{viewlod2_small}~~~~~
    \includegraphics[width=3cm]{subdivision}
  \end{center}
  \caption{Notations \emph{(left)} and adaptive meshing in dimension 2 using the LoD distance and angular criterion \emph{(middle)}. Based on LoD criterion
    evaluation, the cell buffer is updated between each frame \emph{(right)}.}
  \label{fig_lod_octree}
\end{figure}
%
In addition to the distance criterion, we add an angular term in order
to prevent subdivision of cells far from the camera frustrum (see
Figure~\ref{fig_lod_octree}-\emph{middle}). We do not go into the
details but please note that stay, split and merge decisions of all cells are computed in
parallel from all cell morton codes.

@@

\begin{figure}
  \begin{center}
    {\includegraphics[width=2.7cm]{figs/Octa_mean}}
    {\includegraphics[width=2.7cm]{figs/Octa_gaussian}}
    {\includegraphics[width=2.7cm]{figs/Octa_k1}}
    {\includegraphics[width=2.7cm]{figs/Octa_k2}}\\
    {\includegraphics[width=2.7cm]{figs/Octa_dir_min}}
    {\includegraphics[width=2.7cm]{figs/Octa_dir_max}}
    {\includegraphics[height=2.0cm]{figs/Octa_dir_min_zoom}}
    {\includegraphics[height=2.0cm]{figs/Octa_dir_max_zoom}}\\
    % {\includegraphics[height=4.0cm]{figs/EG_dragon_normal_zoom}}
    {\includegraphics[width=2.2cm]{figs/function_mean_0}}
    {\includegraphics[width=2.2cm]{figs/function_mean_1}}
    {\includegraphics[width=2.2cm]{figs/function_mean_2}}
    {\includegraphics[width=2.2cm]{figs/function_mean_3}}
    {\includegraphics[width=2.2cm]{figs/function_mean_4}}
  \end{center}
  \caption{\emph{Third row:} Mean, Gaussian, first and second principal curvature estimation on an OctaFlower with a digital domain of $130^3$.
  \emph{Second row:} First and second principal directions estimation (and zoom) on an OctaFlower.
  \emph{Third row:} Mean curvature computed in real-time on a dynamic object.}
  \label{fig:full}
\end{figure}

We first evaluate the curvature estimation on a full-resolution geometry
obtained by disabling the LoD criterion at the mesh generation step.
Figure~\ref{fig:full}-top and middle shows results of curvature tensor
estimation (mean, Gaussian, first and second principal curvatures, first and
second principal directions) on an OctaFlower at level $l_0$ (ground truth).
Figure~\ref{fig:full}-bottom shows mean curvature estimation in real-time on a
dynamic object. For all illustrations, we use directly normal vectors computed
by algorithm discussed in Section~\ref{sec:curv-tens-estim}.

@@

  \end{center}
  \caption{Comparison of number of texture fetches, $L_2$ and $L_\infty$ error obtained on OctaFlower with a digital domain of $130^3$. We measured it using radii of 8 and 16.
  The object was triangulated at full resolution with a regular grid and contains 282396 vertices. The timings
    only consider the curvature estimation step.\label{tab:full-res-stat}}
\end{table}

@@

\begin{figure}[!htbp]
  \begin{center}
    \begin{tikzpicture}[scale=0.7]
      \begin{axis}[
          ybar stacked,
          ymode = log,
          ymax = 6000,
	  	  bar width=20pt,
	  	  nodes near coords,
	  	  every node near coord/.style={ font=\tiny, inner sep=0pt},
          enlargelimits=0.15,
          ylabel={\#Time (in ms)},
          symbolic x coords={Octa-R8, Octa-R16, Snow-R10, Snow-R20,
	      Dragon-R8, Dragon-R20},
          xtick=data,
          point meta=rawy,
          x tick label style={rotate=45,anchor=east},
        ]
        %Geometry Generation
        \addplot+[ybar, color=black, fill=cgeometry] plot coordinates {(Octa-R8,17.283000) (Octa-R16,18.458000) 
          (Snow-R10,9.831000) (Snow-R20,9.653000)  (Dragon-R8,136.300000) (Dragon-R20,145.450000)};
        %l4
        \addplot+[ybar, color=black, fill=clquatre] plot coordinates {(Octa-R8,1) (Octa-R16,2.509000) 
          (Snow-R10,6.085000) (Snow-R20,9.445000)  (Dragon-R8,3.389000) (Dragon-R20,20.228000)};
        %l3
        \addplot+[ybar, color=black, fill=cltrois] plot coordinates {(Octa-R8,2.527000) (Octa-R16,4.159000)
          (Snow-R10,10.629000) (Snow-R20,11.542000) (Dragon-R8,20.397000) (Dragon-R20,26.710000)};
        %l2
        \addplot+[ybar, color=black, fill=cldeux] plot coordinates {(Octa-R8,3.783000) (Octa-R16,9.006000) 
          (Snow-R10,13.777000) (Snow-R20,37.700000) (Dragon-R8,29.967000) (Dragon-R20,92.435000)};
        %l1
        \addplot+[ybar, color=black, fill=clun] plot coordinates {(Octa-R8, 11.004000) (Octa-R16,38.687000) 
          (Snow-R10,39.627000) (Snow-R20,217.954000) (Dragon-R8,59.840000) (Dragon-R20,506.549000)};
        %l0
        \addplot+[ybar, color=black, fill=clzero] plot coordinates {(Octa-R8, 37.273000) (Octa-R16,244.353000) 
          (Snow-R10,236.194000) (Snow-R20,1655.680000) (Dragon-R8,322.674000) (Dragon-R20,3763.422000)};
        %\legend{\strut Geometry, \strut Curvature, \strut xx, \strut xx}
      \end{axis}
	  \end{tikzpicture}
%
    \begin{tikzpicture}[scale=0.7]
      \begin{axis}[
          ybar stacked,
          ymax = 6000,
          ymode = log,
	  	  bar width=20pt,
	  	  nodes near coords,
	  	  every node near coord/.style={ at={(0,0)}, font=\tiny, inner sep=0pt},
          enlargelimits=0.15,
          %ylabel={\#Time (in ms)},
          symbolic x coords={Octa-R8, Octa-R16, Snow-R10, Snow-R20, 
	      Dragon-R8, Dragon-R20},
          xtick=data,
          point meta=rawy,
          x tick label style={rotate=45,anchor=east},
        ]
        %Geometry Generation
        \addplot+[ybar,color=black, fill=cgeometry] plot coordinates {(Octa-R8,17.283000) (Octa-R16,17.183000) 
          (Snow-R10,9.383000) (Snow-R20,9.767000)  (Dragon-R8,127.347000) (Dragon-R20,127.166000)};
        %Hierarchical measurement
        \addplot+[ybar, color=black, fill=clzero] plot coordinates {(Octa-R8,18.095000) (Octa-R16,64.186000) 
          (Snow-R10,96.747000) (Snow-R20,364.547000)  (Dragon-R8,147.662000) (Dragon-R20,787.459000)};
        %\legend{\strut Geometry, \strut Curvature, \strut xx, \strut xx}
      \end{axis}
     \end{tikzpicture}
     \end{center}
  %
  \caption{Timings in milliseconds obtained while visualizing an adaptive
  triangulation on three objects -- OctaFlower (digital domain of $130^3$ ), Snow
  Microstructures ($233^3$) and XYZ-Dragon ($510^3$) --  by computing the
  curvature with a regular grid (\emph{left}) and with hierarchical
  algorithm (\emph{right}), with two different radii for each object.
  %
  In \emph{orange color}: time required to extract the triangulation.
  %
  In \emph{blue color}: time required to compute the curvature tensor at
  different levels: from $l_4$ (light blue) to $l_0$ (dark blue).
  %
  }
  %
  \label{fig:timings}
\end{figure}

In this experiments, we measure the timings obtained when running our framework on a
GeForce GTX 850M. Here, the triangulation is dynamically refining according to its distance to the camera.
At each frame, both the curvature and the triangulation are recomputed.
If the visualisation stays unchanged, we compute the curvature at a level $l-1$ until the level $l_0$ is reached.
We measured those timings using the regular and the hierarchical algorithm. The results are presented in 
logarithmic scale figure \ref{fig:timings}.

We can first note that the required time for an object to compute ground truth curvature is usually
as high as the time required to extract the geometry and compute all the above levels.
Using approximations is thus mandatory, since for big objects such as the XYZ Dragon (Digital Domain $510^3$)
with high radii, the computation is seconds faster, the $l_0$ visualisation takes more than 4 seconds to compute, whereas
the approximated visualisation at $l_2$ takes only 300ms.
We can thus get an approximated visualisation of our object as soon as we run the application, allowing for
real time interactions with the object (changing the visualised value, the curvature radius, etc ...).
Once an interesting portion if the object is reached, it only takes a few seconds to the application to display
ground truth curvature estimation.

It is also clear on those graphs that using a hierarchical decomposition reduces the curvature computation time.
For the XYZ Dragon with a radius of 20, it needs almost 4 seconds to compute the $l_0$ level, whereas using the
hierarchical algorithm, it only needs 700ms. However, due to the precomputation, this algorithm is biased and
creates an error (presented in Table~\ref{tab:full-res-stat}) that needs to be considered

%\subsection{Visualisation}
%
%For dynamic datasets, we consider a time varying implicit
%surface. Since it would have been to expensive to update the a mipmap texture
%of densities at each time step, we simply evaluate the implicit
%expression at each vertex of a cell to decide if such cell is included
%or not into $\Ball{R}{\vx}\cap \Shape$.
%\todo[inline]{blabla experiments}

@@
